Sophize Mathematics Library (Work In Progress)

### 2022-05-25 math_library/sophize_math.tex
minor
For CICM '22 submission -1/X

\title{Sophize Markdown and Collaboration Interface }

@@

\begin{abstract}
[FROM CICM] Sophize is a novel mathematics library and discussion platform with a mission to help our users find and organize mathematical proofs. We present a Sophize's novel yet familiar knowledge organization scheme that is used to represent a wide variety proofs. With this knowledge organization scheme at its core, we have engineered that platform to aggregate knowledge from sources. The platform meaningfully combines knowledge from published research, encyclopedias such as PlanetMath and Wikipedia, informal computer programs and formal systems such as Metamath; and allows its users to run a federated search over them. To make this happen, two developments were necessary. First, we developed an open plugin architecture that allows proofs from computer programs to be generated as required. Second, we extended the Markdown language to represent the connections between between mathematical objects found across various sources of knowledge. In addition, we also utilized this new language to create a novel communication system built specifically to aid mathematicians in solving problems collaboratively.

@@

\subsection*{Acknowledgements [FROM CICM]}
This publication resulted (in part) from research supported by the International Mathematical Union. The author would like to thank Patrick Ion and Michael Kohlhase for help in organizing our ideas and structuring the content of the paper.

\section{Sophize's Knowledge Organization Scheme}
For math, logical consistency is the sole criteria for truth. Thus, as a math library, Sophize models mathematical knowledge using logical arguments.

\subsection{Core concepts [FROM CICM]}
The various concepts used for organizing knowledge logically are described below. These concepts have been modelled in JSON schema \cite{sophize_datamodel} and published for popular programming language such as Java (MVN), Javascript (npm) and Python (PyPI).

@@

\paragraph{A \textbf{resource} is an abstract concept inherited by all other top-level concepts like terms, propositions, and arguments. Each resource has a URI and contains fields such as search tags and citations.}

There are a couple more concepts that we will describe at later in the paper.

@@

\subsection{Argument Graph}
A proposition is considered true if it is claimed by a sound argument. An argument is considered sound if it is valid and its premises are true. Also, as explained in the section above, the validity criteria of an argument is modelled as just another premise. Thus, the relationship between all propositions and arguments can be represented with a simple graph. The following figure shows the different types of nodes and edges in such a graph

@@

Conjectural knowledge (Reimann hypothesis). 
Practical utility of conjectural knowledge - eg. a lot of infromation security systems assume  P not NP)

>>>>

\title{Sophize Mathematics Library (Work In Progress) }

@@

\begin{abstract}
Sophize is a novel mathematics library and discussion platform with a mission to help our users find and organize mathematical proofs. We present Sophize's novel yet familiar knowledge organization scheme that is used to represent a wide variety proofs. With this knowledge organization scheme at its core, we have engineered a platform to aggregate knowledge from multiple sources. The platform meaningfully combines knowledge from published research, encyclopedias such as PlanetMath and Wikipedia, informal computer programs and formal systems such as Metamath; and allows its users to run a federated search over them. To make this happen, two developments were necessary. First, we developed an open plugin architecture that allows proofs from computer programs to be generated as required. Second, we extended the Markdown language to represent the connections between between mathematical objects found across various sources of knowledge. In addition, we also utilized this new language to create a novel communication system built specifically to aid mathematicians in solving problems collaboratively.

@@

\section{Sophize's Knowledge Organization Scheme}
At its core, Sophize's approach is to simply keep track of all arguments in favour or against any proposition. Use of arguments is the common factor amongst all mathematical (and scientific) knowledge. We use this fact to unify knowledge from multiple sources in a meaningful way as described below.

We use the set of all known arguments to track knowledge from a variety of foundational theory. For theoretical knowledge, i.e., knowledge that doesn't involve empirical data, the primary criteria of validity is internal (logical) consistency. Thus, for each foundational theory (for every belief set to be precise), we also keep track of any contradictions that arise from those foundations.

The following subsections will introduce concepts that will be

\subsection{Core concepts}
The various concepts used for organizing knowledge logically are described below. These concepts have been modelled in JSON schema \cite{sophize_datamodel} and published for popular programming language such as Java (MVN), Javascript (npm) and Python (PyPI).

\paragraph{A \textbf{resource} is an abstract concept inherited by all other top-level concepts like terms, propositions, arguments, and belief sets. Each resource has a URI and contains fields such as search tags and citations.}

@@

Sophize's notions of \textbf{belief sets} and \textbf{proof-generating machine} is described later in this section.

\subsection{Argument Graph}
An argument can be seen as a simple graph (see Figure \ref{argument}) with two types of nodes - (a) a single node representing the argument itself and (b) a set of proposition nodes representing the argument's conclusion and premises. The edge of this graph are directed and go from (a) premise nodes to argument and (b) argument node to conclusion node.

\begin{figure}[ht]
\begin{center}
\fbox{\includegraphics[height=4cm]{argument}}
\caption{An argument represented as a graph.}
\label{argument}
\end{center}
\end{figure}

A set of arguments will can thus be represented as a single graph that we call an \textbf{argument graph}.

\begin{figure}[ht]
\begin{center}
\fbox{\includegraphics[height=7cm]{argument_graph}}
\caption{A set of arguments forming an argument graph.}
\label{argument_graph}
\end{center}
\end{figure}

A proposition is considered true if it is claimed by a sound argument. An argument is considered sound if it is valid and its premises are true. Also, as explained in the section above, the validity criteria of an argument is modelled as just another premise. Thus, the relationship between all propositions and arguments can be represented with a simple graph. The following figure shows the different types of nodes and edges in such a graph

@@

@@

Conjectural knowledge (Reimann hypothesis).

Practical utility of conjectural knowledge - eg. a lot of infromation security systems assume  P not NP)

Proof tailored to needs. Proof graph can be easily converted to prose too.
